---
title: "Project 2"
author: "Michael Lee"
date: "December 19, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#setup
We clean the data and import the necessary packages below. The data we use is the Monthly Price Index from 2012-2018  
https://data.bls.gov/cgi-bin/surveymost
```{r}
library(readxl)
library(ggplot2)
library(e1071)
library(goftest)
library(tseries)
library(forecast)

importRaw <- read_excel("C:/Users/micha/OneDrive/Documents/2018-2019/Time Series/Import.xlsx")

#trims the raw data for use
importRaw <- importRaw[11:length(importRaw$X__1),]
importRaw$`Import/Export Price Indexes` <- NULL

#this loop converts the CPI prices into a single variable
index <- c()
for(i in c(1:length(importRaw$X__1))){
  a <- as.numeric(importRaw[i,])
  a <- a[-1]
  index <- append(index, a)
}
import <- data.frame(index)
import <- na.omit(import)

#creates a time variable representative of each month 
import$time <- c(1:length(import$index))
import <- import[import$time >35,]
import$time <- c(1:length(import$index))
import <- import[,c(2,1)]

```


#Initial Analysis

The initial plot shows us that there is level dependent volatility. 
```{r}
ggplot(import, aes(x = time, y = index)) +
  geom_line(color = "#00AFBB", size = 1) +
  labs(x="Time", y="Index", title="Import Price Index") +
  theme_bw()
```

We take the logs of our data and save the last observation (log85) to evaluate our intervals later.  
The ACF and PACF indicate we should difference our data. 
```{r}
import$logIndex <- log(import$index)
log85 <- import$logIndex[85]
import <- import[-c(length(import$time)),]
import$logDiff <- c(NA,diff(import$logIndex))
```

```{r}
ggplot(import, aes(x = time, y = logIndex)) +
  geom_line(color = "#00AFBB", size = 1) +
  labs(x="Time", y="Index", title="Log of Index") +
  theme_bw()
```

```{r}
acf(import$logIndex)
```



```{r}
pacf(import$logIndex)
```



```{r}
ggplot(import, aes(x = time, y = logDiff)) +
  geom_line(color = "#00AFBB", size = 1) +
  labs(x="Time", y="Index", title="Log Difference of Index") +
  theme_bw()
```

```{r}
acf(na.omit(import$logDiff))
```

```{r}
pacf(na.omit(import$logDiff))
```

#Selecting a Model

Below we examine the AICC for various ARIMA models using d = 1. AICC suggests a ARIMA(0,1,2) model
```{r}
#with constants
indexTS <- ts(import$logIndex, start = 1, end = length(import$logIndex))

x <- c(0,1,2,3)
z <- c(0,1,2,3)
#with constants
results <- c()
pdq <- c()
for(i in x){
  for(j in z){
    model <- Arima(indexTS, c(i,1,j), include.constant = TRUE)
    results <- c(results, model$aic)
    pdq <-c(pdq, paste(i,"1",j, sep= " "))
  }
}
AIC1 <- data.frame(pdq, results)
print(AIC1)
```

```{r}
#without constants
results <- c()
pdq <- c()
for(i in x){
  for(j in z){
    model <- Arima(indexTS, c(i,1,j), include.constant = FALSE)
    results <- c(results, model$aic)
    pdq <-c(pdq, paste(i,"1",j, sep= " "))
  }
}
AIC2 <- data.frame(pdq, results)
print(AIC2)
#ARIMA(0,1,2)
```


```{r}
arimaModel <- Arima(indexTS,c(0,1,2),include.constant = FALSE)
arimaModel
```

The Model is:
$$
\ X_{t} = \epsilon_t - .6192\epsilon_{t-1} - .1043\epsilon_{t-2} 
$$


Our forecast for the one step ahead interval does not include our actual data point, 4.8378 meaning our intervals are not wide enough. 
```{r}
forecast <- forecast(arimaModel, h = 1, level = 95)
forecastData<- data.frame(forecast)
forecastData

```


#Residuals

The plots below show that the residuals are not normally distributed and there is statistically significant autocorrelation   
of the squared residuals. This means our data has conditional heteroskedasticity and an ARCH model is appropriate.s



plot of residuals
```{r}
import$resids <- arimaModel$residuals
import$residsSQR <- (import$resids)^2

plot(import$resids)
```

plot of residuals squared
```{r}
plot(import$residsSQR)
```

acf of residsuals
```{r}
acf(import$resids)
```

pacf of residuals
```{r}
pacf(import$resids)
```

acf of residuals squared
```{r}
acf(import$residsSQR)
```

pacf of residuals squared
```{r}
pacf(import$residsSQR)
```

#ARCH Model
Judging from the tables bleow, the GARCH(1,1) has the best AICC so we will select this model.  

AICC and log liklihood for the ARCH(q) Models
```{r}

loglik <- c()
N <- length(import$resids)
for(i in 0:10){
  if(i==0){
    loglik <- c(loglik,-0.5 * N * (1 + log(2 * pi * mean(import$resids^2))))
  }
  else{
    fit <- garch(import$resids, c(0,i), trace=FALSE)
    loglik <- c(loglik, logLik(fit))
  }
}

q <- 0
AICC3 <- c()
for(i in loglik){
  k <- q + 1
  AICC3 <- c(AICC3, (-2 * loglik[q+1]  + 2 * k * N / (N - k - 1)))
  q <- q+1
}
q <- c(0:10)
selection <- data.frame(q, loglik, AICC3)
selection
```

AICC and Log Liklihood for the GARCH(1,1) Model

```{r}
fit <- garch(import$resids, c(1,1), trace=FALSE)
loglik <- logLik(fit)
k <- 2
aicc <- -2 * loglik  + 2 * k * N / (N - k - 1)
print(data.frame(loglik, aicc))
```


```{r}
summary(fit)
```

The Model is:

$$
\sigma^{2}_{t} = 2.46\times10^{-5}+0.3024\epsilon_{t-1}^{2} + 0.2519\sigma_{t-1}^{2}
$$

unconditional variance
```{r}
uv <- (2.460 * 10^(-05))/(1-.3024-.252)
uv
```

#Forecast Intervals

This shows the lower and upper bound of our forecast. They seem to be higher than our intervals from the Arima Model.
```{r}
f1 <- fitted.values(arimaModel)
h1 <- fit$fit[,1]^2
lower <- f1 + -1.96 * sqrt(h1)
upper <- f1 + 1.96 * sqrt(h1)
data.frame(lower,upper)

```


This plot shows that there are high bursts of volatility. 
```{r}
tail(h1,n=10)
plot(import$time, h1, type="l", col=4)
```

The forecast intervals we generated follow the data well. 
```{r}
ggplot() +
  geom_line(aes(import$time, import$logIndex)) +
  geom_line(aes(import$time, f1 + 1.96*sqrt(h1), col = "High 95%")) +
  geom_line(aes(import$time, f1 - 1.96*sqrt(h1), col = "Low 95%")) +
  labs(x = "Months", y = "Logged Rupee") +
  theme_bw()

```



#Arch Residuals

The probability plot shows that our ARCH Residuals are approximatly normal. 
```{r}

archResid <- import$resids / sqrt(h1)
probplot(archResid)
```

This shows that our forecast intervals failed 3.57% of the time
```{r}
failed <- sum(abs(import$logIndex - f1) > sqrt(h1) * 1.96, na.rm=TRUE)/nrow(import)
failed
```
